---
title: "Views"
author: "Bernardo Reckziegel"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Views}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- As mentioned in __How does EP work__, entropy-pooling offers a template that enables portfolio and risk managers to process _views_ that can go far beyond normality.  -->

This vignette provides a "jump-start" for users that want to quickly understand how to use entropy-pooling (EP) to construct *views* on the market distribution. The methodology fits well in a variety of scenarios: stress-testing, macro factors, non-linear views, and more. 

To demonstrate entropy-pooling's firepower, the current vignette uses some of the cases Meucci (2008)[^1] leaves for the reader in the appendix. In particular, the usage of the following functions are covered in depth:

- `view_on_mean()`
- `view_on_volatility()`
- `view_on_correlation()`
- `view_on_rank()`
- `view_on_marginal_distribution()`
- `view_on_copula()`

[^1]: Meucci, Attilio, Fully Flexible Views: Theory and Practice (August 8, 2008). Fully Flexible Views: Theory and Practice, Risk, Vol. 21, No. 10, pp. 97-102, October 2008, Available at SSRN: https://ssrn.com/abstract=1213325

Along with the examples, the `EuStockMarkets` dataset - that comes with the installation of `R` - is used as a _proxy_ for "the market":

```{r setup}
library(ffp)
library(ggplot2)
set.seed(321)

# invariance / stationarity
x <- diff(log(EuStockMarkets))

dim(x)

head(x)
```

<!-- The object `x` is a $1859 \times 4$ matrix with four European stock indices (`DAX`, `SMI`, `CAC` and `FTSE`) that are sampled from $1991$ to $1998$.  -->

<!-- ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=7, fig.align='center'} -->
<!-- library(ggplot2) -->
<!-- tibble::as_tibble(x) |>  -->
<!--   tidyr::pivot_longer(cols = dplyr::everything()) |>  -->
<!--   ggplot(aes(x = value, fill = name, color = name)) +  -->
<!--   geom_density(show.legend = FALSE) +  -->
<!--   facet_wrap(~name) +  -->
<!--   scale_color_viridis_d(end = 0.75, option = "C") +  -->
<!--   scale_fill_viridis_d(end = 0.75, option = "C") +  -->
<!--   scale_x_continuous(labels = scales::percent_format(), limits = c(-0.05, 0.05)) + -->
<!--   scale_y_continuous(labels = scales::percent_format(scale = 1)) +  -->
<!--   labs(title = "Marginal Distributions",  -->
<!--     subtitle = "European Stock Indexes", -->
<!--     x = NULL, y = NULL, color = NULL, fill = NULL) -->
<!-- ``` -->

## Views on Expected Returns

Assume an investor believes the `FTSE` index return will be $20\%$ above average in the near future. In order to process this *view*, the investor needs to minimize the relative entropy:

$$ \sum_{j=1}^J x_j(ln(x_j) - ln(p_j)) $$ 

Subject to the restriction:

$$ \ x_j V_{j, k} = \mu_{k} $$

In which $x_{j}$ is a yet to be discovered _posterior_ probability; $V_{j,k}$ is a $1859 \times 1$ vector with the `FTSE` returns; and $\mu_{k}$ is a $1 \times 1$ scalar with the investor projected return for the `FTSE` index. In this case, the $k$ subscript represents the fourth column in `x`[^2].

[^2]: The commands `x[ , 4]` or `x[ , "FTSE"]` yield the same results.

*Views* on expected returns can be constructed with `view_on_mean`:

```{r}
# Returns 20% higher than average
mu <- mean(x[ , "FTSE"]) * 1.2

# ffp views constructor
views <- view_on_mean(x = as.matrix(x[ , "FTSE"]), mean = mu)
views
```

The `views` object is a `list` with two components, `Aeq` and `beq`, that are equivalent to the elements $V_{j,k}$ transposed and $\mu_{k}$, respectively.

The investor also needs to formulate a vector of *prior* probabilities, $p_j$, which is usually - but not necessary - a equal-weight vector:

```{r}
# Prior probabilities
p_j <- rep(1 / nrow(x), nrow(x)) 
```

Once the _prior_ and the _views_ are established, the optimization can take place with `entropy_pooling`:

```{r}
# Solve Minimum Relative Entropy (MRE)
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nlminb"
)
ep
```

What is the interpretation for `ep`? Among all the possible probability vectors, `ep` is the one that can satisfy the *views* in the "least" evasive way (in which the term "least evasive" is direct reference to the _prior_). Therefore, `ep` is the best candidate for a _posterior_ distribution. 

<!-- The main advantage of using the _views_ constructed in the `ffp` package is that, as a user, you just have to pass the arguments thorough the `entropy_pooling` function. -->

In real world, the investor can have many *views*. Extending the current example, say the investor has a new belief: the `CAC` index returns will be $10\%$ bellow average[^3]:

[^3]: An ARIMA, GARCH or VAR model could be used as well. The focus is not on the best forecasting method, but how to input the _view_ once the forecast has already been made.

```{r}
mu <- c(
  mean(x[ , "FTSE"]) * 1.2, # Return 20% higher than average
  mean(x[ , "CAC"]) * 0.9   # Return 10% bellow average
)

# ffp views constructor
views <- view_on_mean(x = as.matrix(x[ , c("FTSE", "CAC")]), mean = mu)
views
```

In which the elements `Aeq` and `beq` now have $2$ rows each, one for every _view_. 

The minimum relative entropy problem is solved in the exact same way:

```{r}
# Solve MRE
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nlminb"
)
ep
```

It's easier to analyse the output of `ep` visually. 

```{r}
class(ep)
```

To that end, the `autoplot` method is available for objects of the `ffp` class:

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "Views on Expected Returns", 
       x        = NULL, 
       y        = NULL)
```

The plot reinforces the intuition of $x_j$ - the _posterior_ probability - as a distribution that distorts the _prior_ in order to accommodate the _views_.

It's easy to double-check the investor *views* by computing the ratio of the *conditional* vs. *unconditional* returns:

```{r}
conditional   <- ffp_moments(x, ep)$mu 
unconditional <- colMeans(x)

conditional / unconditional - 1
```

According to expectations, `CAC` outlook is now $10\%$ bellow average and `FTSE` expected return is $20\%$ above average. _Violà!_

## Views on Volatility

Say the investor is confident that markets will be followed by a period of lull and bonanza. Without additional insights about expected returns, the investor anticipates that market volatility for the `CAC` index will drop by $10\%$. 

To impose _views_ on volatility, the investor needs to minimize the expression:

$$ \sum_{j=1}^J x_j(ln(x_j) - ln(p_j)) $$
Subject to the constraint: 

$$ \sum_{j=1}^{J} x_j V_{j, k}^2 = \mu_{k}^2 + \sigma_{k}^2 $$

In which, $x_j$ is a vector of yet to be defined _posterior_ probabilities; $V_{j,k}^2$ is a $1859 \times 1$ vector with `CAC` squared returns; $\mu_{k}^2$ and $\sigma_{k}^2$ are the `CAC` sample mean and variance; and $k$ is the column position of `CAC` in the panel $V$. 

The `view_on_volatility` function can be used for this purpose: 

```{r}
# opinion
vol_cond <- sd(x[ , "CAC"]) * 0.9

# views
views <- view_on_volatility(x = as.matrix(x[ , "CAC"]), vol = vol_cond)
views
```

The `views` object holds a list with two components - `Aeq` and `beq` - that are equivalent to the elements $V_{j,k}^2$ transposed and $\mu_{k}^2 + \sigma_{k}^2$, respectively. 

To solve the relative entropy use the `entropy_pooling` function:

```{r}
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nlminb"
)
ep
```

Once again, the `ep` vector is what we call *posterior*: the probability vector that causes the "least" amount of distortion in the *prior*, but still obeys the constraints (*views*).

<!-- Notice that the same commands are used to find a solution in different problems (expected returns and volatilities). That's exactly the objective of `view_on_*()` family of functions: to provide a template for entropy minimization! -->

The *posterior* probabilities can be observed with the `autoplot` method:

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "View on Volatility", 
       x        = NULL, 
       y        = NULL)
```

To check if the _posterior_ probabilities are valid as a way to display the investor's _view_, compare the _conditional_ vs. _unconditional_ volatilities:  

```{r}
unconditional <- apply(x, 2, sd)
conditional   <- sqrt(diag(ffp_moments(x, ep)$sigma))

conditional / unconditional - 1
```

The _posterior_ volatility for the `CAC` index is reduced by $10\%$, in line with the investor's subjective judgment. However, by emphasizing tranquil periods more often, other assets are also affected, but in smaller magnitudes.

## Views on Correlation

Assume the investor believes the correlation between `FTSE` and `DAX` will increase by $30\%$, from $0.64$ to $0.83$. To construct *views* on the correlation matrix, the general expression has to be minimized:

$$ \sum_{j=1}^J x_j(ln(x_j) - ln(p_j)) $$
Subject to the constraints: 

$$ \sum_{j=1}^{J} x_j V_{j, k} V_{j, l} = \mu_{k} \mu_{l} + \sigma_{k} \sigma_{l} C_{k,l} $$

In which, the term $V_{j, k} V_{j, l}$ on the left hand-side of the restriction consists of cross-correlations among assets; the terms $\mu_{k} \mu_{l} + \sigma_{k} \sigma_{l} C_{k,l}$ carry the investor target correlation structure; and $x_j$ is a yet to be defined probability vector.

To formulate this *view*, first compute the _unconditional_ correlation matrix. Second, add a "perturbation" in the corresponding element that is consistent with the perceived *view*:

```{r}
cor_view <- cor(x) # unconditional correlation matrix
cor_view["FTSE", "DAX"] <- 0.83 # perturbation (investor belief)
cor_view["DAX", "FTSE"] <- 0.83 # perturbation (investor belief)
```

Finally, pass the adjusted correlation matrix into the `view_on_correlation` function:

```{r}
views <- view_on_correlation(x = x, cor = cor_view)
views
```

In which `Aeq` is equal to $V_{j, k} V_{j, l}$ transposed and `beq` is a $10 \times 1$ vector with $\mu_{k} \mu_{l} + \sigma_{k} \sigma_{l} C_{k,l}$.

Notice that even though the investor has a *view* in just one parameter[^4] the constraint vector has $10$ rows, because every element of the lower/upper diagonal of the correlation matrix has to match.

[^4]: `cor(ftse, dax) = 0.83`.

Once again, the minimum entropy is solved by `entropy_pooling`:

```{r}
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nlminb"
)
ep
```

And the `autoplot` method is available:

<!-- so that the *posterior* probabilities can be inspected more easily: -->

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "View on Correlation", 
       x        = NULL, 
       y        = NULL)
```

To fact-check the investor _view_, compute the *posterior* correlation matrix:

```{r}
cov2cor(ffp_moments(x, p = ep)$sigma)
```

And notice that the linear association between `FTSE` and `DAX` is now $0.83$!

## Views on Relative Performance

Assume the investor believes the `DAX` index will outperform the `SMI` by some amount, but he doesn't know by how much. If the investor has only a mild *view* on the performance of assets, he may want to minimize the following expression:

$$ \sum_{j=1}^J x_j(ln(x_j) - ln(p_j)) $$
Subject to the constraint:

$$ \sum_{j=1}^{J} x_j (V_{j, k} - V_{j, l}) \ge 0 $$
In which, the $x_j$ is a yet to be determined probability vector; $V_{j, k}$ is a $1859 \times 1$ column vector with `DAX` returns and $V_{j, l}$ is a $1859 \times 1$ column vector with the `SMI` returns. In this case, the $k$ and $l$ subscripts refers to the column positions of `DAX` and `SMI` in the object `x`, respectively.

_Views_ on relative performance can be imposed with `view_on_rank`:

```{r}
views <- view_on_rank(x = x, rank = c(2, 1))
views
```

Assets that are expected to outperform enter in the `rank` argument with their column positions to the right: assets in the left underperform and assets in the right outperform. Because the investor believes that `DAX` $\geq$ `SMI` (the asset in the first column will outperform the asset in the second column), we fill `rank = c(2, 1)`[^5].

[^5]: In the `x` object, `DAX` it's in the first column and the `SMI` it's in the second column.

The optimization is, once again, guided by the `entropy_pooling` function:

```{r}
ep <- entropy_pooling(
  p      = p_j, 
  A      = views$A, 
  b      = views$b, 
  solver = "nloptr"
)
ep
```

Two important observations:

- Inequalities constraint cannot be handled with the `nlminb` solver. Use `nloptr` or `solnl`, instead;
- Inequalities constraint require the arguments `A` and `b` to be fulfilled rather than `Aeq` and `beq`.

The _posterior_ probabilities are presented bellow:

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "View on Ranking/Momentum", 
       x        = NULL, 
       y        = NULL)
```

To fact-check the ongoing *view*, compare the *conditional* vs. *unconditional* returns:

```{r}
conditional   <- ffp_moments(x, ep)$mu  
unconditional <- colMeans(x)

conditional / unconditional - 1
```

Indeed, returns for `DAX` are adjusted upwards by $17\%$, while returns for `SMI` are revised downwards by $6.5\%$. Notice that returns for `CAC` and `FTSE` were also affected. This behavior could be tamed by combining the ranking condition with _views_ on expected returns. See the function `bind_views()` for an example. 

## Views on Marginal Distribution

Assume the investor has a _view_ on the entire marginal distribution. One way to add _views_ on marginal distributions is by matching the first moments exactly, up to a given order. Mathematically, the investor minimizes the relative entropy: 

$$ \sum_{j=1}^J x_j(ln(x_j) - ln(p_j)) $$
With, say, two equality constraints (one for $\mu$ and one for $\sigma^2$):

$$ \sum_{j=1}^J x_j V_{j,k}  =  \sum_{j=z}^Z p_z \hat{V}_{z,k} $$
$$ \sum_{j=1}^J x_j (V_{j,k})^2  =  \sum_{z=1}^Z p_z (\hat{V}_{z,k})^2 $$
<!-- $$ \sum_{j=1}^J x_j (V_{j,k})^3  =  \sum_{z=1}^Z p_z (\hat{V}_{z,k})^3 $$ -->
<!-- $$ \sum_{j=1}^J x_j (V_{j,k})^4  =  \sum_{z=1}^Z p_z (\hat{V}_{z,k})^4 $$ -->

In which $x_j$ is a yet to be defined probability vector; $V_{j,k}$ is a matrix with the empirical marginal distributions; $p_z$ is vector of _prior_ probabilities; and $\hat{V}_{z,k}$ an exogenous panel with simulations that are consistent with the investor's _views_. 

When $j = z$, the panels $V_{j,k}$ and $\hat{V}_{z,k}$ have the same number of rows and the dimensions in both sides of the restrictions match. However, it's possible to set $z \ge j$ to simulate a larger panel for $\hat{V}_{z, k}$. Keep in mind though that, if $z \ne j$, two _prior_ probabilities will have to be specified: one for $p_j$ (the objective function) and one for $p_z$ (the _views_). 

Continuing on the example, consider the margins of `x` can be approximated by a symmetric multivariate t-distribution. If this is the case, the estimation can be conducted by the amazing [ghyp](https://cran.r-project.org/web/packages/ghyp/index.html) package, that covers the entire family of [generalized hyperbolic distributions](https://en.wikipedia.org/wiki/Generalised_hyperbolic_distribution):

<!-- The optimization is assisted by the Expectation Maximization (EM) Algorithm that handle univariate as well as multivariate time series: -->

```{r, message=FALSE, warning=FALSE}
library(ghyp)

# multivariate t distribution
dist_fit <- fit.tmv(data = x, symmetric = TRUE, silent = TRUE)
dist_fit
```

The investor then, can construct a large panel with statistical properties similar to the margins he envisions: 

```{r}
set.seed(123)
# random numbers from `dist_fit`
simul  <- rghyp(n = 100000, object = dist_fit)
```

Where the $100.000 \times 4$ matrix is used to match the _views_ with `view_on_marginal_distribution`: 

```{r}
p_z <- rep(1 / 100000, 100000)
views <- view_on_marginal_distribution(
  x     = x, 
  simul = simul, 
  p     = p_z
)
views
```

The objects `simul` and `p_z` corresponds to the terms $\hat{V}_{z,k}$ transposed and $p_z$, respectively. Note that $8$ restrictions are created, four for each moment of the marginal distribution (there are four assets in "the market").

With the _prior_ and the _views_ at hand, the optimization can be quickly implemented:

```{r}
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nloptr"
)
ep
```

As the visual inspection of the vector `ep`:

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "View on Marginal Distribution", 
       x        = NULL, 
       y        = NULL)
```

It's easy to extend the current example. For instance, the investor could "tweak" some of the fitted parameters for stress-testing (i.e. change the degrees of freedom, increase/decrease expected returns, etc) to verify the immediate impact on the P&L (VaR, CVaR, etc).

Acknowledge that when the restrictions are to harsh, the scenarios will be heavily distorted to satisfy the _views_. In this case, one may need to compute the effective number of scenarios (ENS). This is very easy to do with `ens()`:

```{r}
ens(ep)
```

Rest on the investor to calibrate this statistic in order to get a desired confidence level[^6].

[^6]: Meucci, Attilio, Effective Number of Scenarios in Fully Flexible Probabilities (January 30, 2012). GARP Risk Professional, pp. 32-35, February 2012, Available at SSRN: https://ssrn.com/abstract=1971808.

## Views on Copulas

Assume the investor would like to simulate a different dependence structure for the market. _Views_ that change the interdependence of assets can be implemented by minimizing the relative entropy: 

$$ \sum_{j=1}^J x_j(ln(x_j) - ln(p_j)) $$
Subject to the following restrictions: 

$$ \sum_{j=1}^J x_i U_{j,k} = 0.5$$
<!-- $$ \sum_{j=1}^J x_i U_{j,k}^2 = 0.33 $$ -->
$$ \sum_{j=1}^J x_j U_{j,k}U_{j,l}  =  \sum_{z=1}^Z p_z \hat{U}_{z,k}\hat{U}_{z,l} $$


$$ \sum_{j=1}^J x_j U_{j,k}U_{j,l}U_{j,i}  =  \sum_{j=1}^J p_j \hat{U}_{j,k}\hat{U}_{j,l}\hat{U}_{j,i} $$

In which, the first restriction matches the first moment of the uniform distribution; the second and third restrictions pair the cross-moments of the empirical copula, $U$, with the simulated copula, $\hat{U}$; $x_j$ is a yet to be discovered _posterior_ distribution; $p_z$ is a _prior_ probability; When $j = z$, the dimensions of $p_j$ and $p_z$ match.

Among many of the available copulas, say the investor wants to model the dependence of the market as a [clayton copula](https://en.wikipedia.org/wiki/Copula_(probability_theory)) to ensure the lower [tail dependency](https://en.wikipedia.org/wiki/Tail_dependence) does not go unnoticed. The estimation is simple to implement with the package [copula](https://cran.r-project.org/web/packages/copula/index.html):

```{r, warning=FALSE, message=FALSE}
library(copula)

# copula (pseudo-observation)
u <- pobs(x)

# copula estimation
cop <- fitCopula(
  copula = claytonCopula(dim = ncol(x)), 
  data   = u, 
  method = "ml"
)

# simulation
r_cop <- rCopula(
  n      = 100000, 
  copula = claytonCopula(param = cop@estimate, dim = 4)
)
```

First, the copulas are computed component-wise by applying the marginal empirical distribution in every $k$ column of the dataset. Second, the pseudo-observations are used to fit a Clayton copula by the Maximum-Likelihood (ML) method. Finally, a large panel with dimension $100.000 \times 4$ is constructed to match a Clayton copula with $\hat \alpha = 1.06$, as in the object `cop`. 

The _views_ on the market is constructed with the `view_on_copula`:

```{r}
views <- view_on_copula(x = u, simul = r_cop, p = p_z)
views
```

Once again, the solution of the minimum relative entropy can be found in one line of code:

```{r}
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nloptr"
)
ep
```

And the `autoplot` method is available for the objects of the `ffp` class: 

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "View on Copulas", 
       x        = NULL, 
       y        = NULL) 
```

To stretch the current example, assume the investor stresses the parameter $\hat \alpha = 1.06$[^9]. Higher $\alpha$ values are linked to extreme occurrences on the left tails of the distribution. Whence, consider the case where $\hat \alpha = 5$:

[^9]: `cop@estimate`

```{r}
cop@estimate <- 5
```

Rerun the previous steps (simulation, _views_ and optimization):

```{r}
# simulation
r_cop <- rCopula(
  n      = 100000, 
  copula = claytonCopula(param = cop@estimate, dim = 4)
)

# #views
views <- view_on_copula(x = u, simul = r_cop, p = p_z)

# relative entropy
ep <- entropy_pooling(
  p      = p_j, 
  Aeq    = views$Aeq, 
  beq    = views$beq, 
  solver = "nloptr"
)
```

To find the new _posterior_ probability vector the satisfy the stress-test condition:

```{r, fig.width=7, fig.align='center'}
autoplot(ep) + 
  scale_color_viridis_c(option = "C", end = 0.75) + 
  labs(title    = "Posterior Probability Distribution", 
       subtitle = "View on Copulas", 
       x        = NULL, 
       y        = NULL) 
```

